Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2020-05-10 manuscript/manuscript.tex
editing Ne section

In the linked selection literature, it is common to translate the impact of
linked selection as a reduction in the level of neutral pairwise diversity in
the absence of linked selection, $\pi_0$. Under the coalescent, $\pi_0 = 2 \mu
\E(T_2)$ where $\E(T_2)$ is the pairwise time to coalescence, which under drift
alone in a constant population of size $N$, is $\E(T_2) = 2N$. The fraction of
neutral diversity (in the absence of linked selection) reduced by selection is
then, $1 - \nicefrac{\bar{\pi}}{\pi_0}$, or equivalently, $1 -
\nicefrac{N_e}{N_{e,D}}$. Then,

\begin{align}
  G(t) \lesssim 1 - \frac{N_e}{N_{e,D}} 
\end{align}

which shows that our measure $G(t)$ is a lower bound, over much shorter
timescales, to the familiar measure $1 - \nicefrac{\bar{\pi}}{\pi}$.

Using another approximation, we can consider what roughly equivalently
reduction in diversity due to background selection is expected, using the
theory of Hudson and Kaplan (\citeyear{Hudson1994-oh},
\citeyear{Hudson1995-xc}). Their theory states,

\begin{align}
  \frac{N_e}{N_{e,D}} = \exp\left(-\frac{U}{R}\right)
\end{align}

where $U$ is the total genome-wide deleterious mutation rate per generation,
and $R$ is the total recombination map length in Morgans. For \emph{Drosophila
melanogaster} (we currently lack good estimates for \emph{D. simulans}), $R
\approx 3$ Morgans, implying a $G(t)$ of at least 20\% implies $U \gtrsim
0.67$. This is entirely consistent with recent work estimating $U \approx 1.6$
in \emph{D. melanogaster} from longer timescale population genetic data
\parencite{Elyashiv2016-vt} (their estimate of $U$ implies a $G(t) \approx
0.59$, or around 59\% reduction in neutral diversity

>>>>

In the linked selection literature, it is common to translate the impact of
linked selection as a reduction in the level of neutral pairwise diversity in
the absence of linked selection, $\pi_0$. Under the coalescent, $\pi_0 = 2 \mu
\E(T_2)$ where $\E(T_2)$ is the pairwise time to coalescence, which under drift
alone in a constant population of size $N_e$, is $\E(T_2) = 2N$. The fraction of
neutral diversity (in the absence of linked selection) reduced by selection is
then, $1 - \nicefrac{\bar{\pi}}{\pi_0}$, or equivalently, $1 -
\nicefrac{N_e}{N_{e,D}}$. Then,

\begin{align}
  G(t) \lesssim 1 - \frac{N_e}{N_{e,D}} 
\end{align}
which shows that our measure $G(t)$ is a lower bound, over much shorter
timescales, to the familiar measure $1 -
\nicefrac{\bar{\pi}}{\pi_{\gc{0}}}$.

\gc{\cite{Elyashiv2016-vt} estimated that linked selection in \emph{Drosophila
  melanogaster} had resulted in a $1 -
\nicefrac{\bar{\pi}}{\pi_{\gc{0}}} = 77\%$ reduction in the level of
neutral diversity. Thus our estimate of $G(t) \approx 20\%$ in \emph{Drosophila
  simulans} is smaller
than that seen over long timespans in a closely related
species. }

% Using another approximation, we can consider what roughly equivalently
% reduction in diversity due to background selection is expected, using the
% theory of Hudson and Kaplan (\citeyear{Hudson1994-oh},
% \citeyear{Hudson1995-xc}). Their theory states,

% \begin{align}
%   \frac{N_e}{N_{e,D}} = \exp\left(-\frac{U}{R}\right)
% \end{align}

% where $U$ is the total genome-wide deleterious mutation rate per generation,
% and $R$ is the total recombination map length in Morgans. For \emph{Drosophila
% melanogaster} (we currently lack good estimates for \emph{D. simulans}), $R
% \approx 3$ Morgans, implying a $G(t)$ of at least 20\% implies $U \gtrsim
% 0.67$. This is entirely consistent with recent work estimating $U \approx 1.6$
% in \emph{D. melanogaster} from longer timescale population genetic data
% \parencite{Elyashiv2016-vt} (their estimate of $U$ implies a $G(t) \approx
% 0.59$, or around 59\% reduction in neutral diversity
